---
title: "Homework 3"
author: "Bhavya Deepti Vadavalli"
date: "2023-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries: 

1. Installing and getting needed packes into the environment 
2. Importing the dataset into the environment. 

```{r preliminaries}
library(curl)
library(ggplot2) #I would've just called library(tidyverse) but R is telling me to update and I really don't want to. R updates terrify me.  
library(rmdformats)


f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/zombies.csv") #importing the zombies data set from Chris' github
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

# Assignment problems

## Question 1: Calculating the population means and standard deviations. 

var() and sd() commands as these are for samples. 
Therefore, I made a function stdev() to calculate standard deviation of the population, using population mean and n (from length(d$variable))

```{r problem1}

#writing a function for population standard deviation of variable y 
stdev <- function(y) {
  sqrt(sum((y - mean(y))^2)/length(y))}

#height
mean(d$height) 
stdev(d$height)

#weight
mean(d$weight) 
stdev(d$weight)

#No. of zombies killed
mean(d$zombies_killed) 
stdev(d$zombies_killed)

#Years of education
mean(d$years_of_education)
stdev(d$years_of_education)

#Age
mean(d$age)
stdev(d$age)
```
## Question 2: Use {ggplot} to make boxplots of each of these variables by gender.

```{r problem2}

# Variable = height
p <- ggplot(data = d, aes(x = gender, y = height,
    color = factor(gender))) + geom_boxplot()
p

# Variable = weight
p <- ggplot(data = d, aes(x = gender, y = weight,
    color = factor(gender))) + geom_boxplot()
p

# Variable = age
p <- ggplot(data = d, aes(x = gender, y = age,
    color = factor(gender))) + geom_boxplot()
p

# Variable = No of zombies killed
p <- ggplot(data = d, aes(x = gender, y = zombies_killed,
    color = factor(gender))) + geom_boxplot()
p

# Variable = Years of education
p <- ggplot(data = d, aes(x = gender, y = years_of_education,
    color = factor(gender))) + geom_boxplot()
p
```

## Question 3: Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?

```{r problem3}
#Age v/s height

p <- ggplot(data = d, aes(x = age, y = height)) + geom_point()
p

#Age v/s weight

p <- ggplot(data = d, aes(x = age, y = weight)) + geom_point()
p
```
Yes, there seems to be a positive correlation between age and both height and weight. This means, as age increases, both height and weight are likely to increase as well. 

However, the relationship between age and height seems to be more correlated upto 25 years of age, after which it flattens. 

The relationship between age and weight is much more linear, but seems to have a lot of scatter around where a best-fit regression would be. 
